---
title: "Introduction to Datakindr"
author: "Cormac Nolan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Datakindr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = F, include = F, warning = F, message=F, error=F}

library(ggplot2)
library(knitr)
library(datakindr)

opts_chunk$set(fig.width=6, fig.height=4)
opts_chunk$set(echo = F, warning = F, error = F, message = F)

```
## What Is It For?
Datakindr provides tools, utilities, data and more for use by the DataKind Dublin team. It includes a convenient interface to CSO.ie's Statbank, a ggplot2 theme and colour palettes for convenient & consistent visualisations.

The package is in the early stages of development but it is intended that this package will become a kind of "Swiss Army Knife". It should mean easy and frictionless solving of commonly encountered problems for Datakind volunteers (e.g., querying the CSO's Statbank).

## What Does It Do?  
Currently there are two main tasks that datakindr makes easier:

1. Querying the CSO's Statbank API.
2. Making consistent & beautiful visualisations.

### CSO and Statbank
There are two functions available for this task: 

* `search_statbank_dataset()` Gets a list of datasets from a search query.
* `get_cso_dataset()` Gets the actual data from Statbank.

### Visualisations
There are two sets of data objects here for helping volunteers create consistent plots and other visualisations.

* `dk_theme` A ggplot2 object that will help you effortlessly create visualisations like this:
```{r first plot, echo=FALSE, fig.width = 7, fig.height = 3}
ggplot(data.frame( y = runif(100, min = 0, max = 20) +
                     seq(1, 100, 1),
                   x = seq(1, 100, 1),
                   z = rep_len(c("a", "b", "c", "d"), 100)),
       aes(x, y, colour = z, fill = z)) +
  geom_point() +
  facet_wrap(~z, nrow = 1) +
  geom_smooth(se = TRUE) +
  labs(title = "Some Data (2011)",
       x = "Range", y = "Value") +
  dk_theme


```
* `dk_palette` A palette of 2 primary, 2 secondary and 2 background colours. The primary and secondary colours can be seen above.  
    + `dk_sequential_palette` An extended version of the above palette intended for use with sequential data.
    + `dk_non_sequential_palette` Another extended palette for (obviously!) non-sequential data.

## How Do I Use It?
I'll let the code do the talking here.

### Theme Demo
```{r using dk_theme, echo = T}
# install.packages("devtools")
# devtools::install_github("DataKind-DUB/datakindr")
# Development version: devtools::install_github("cormac85/datakindr")
library(datakindr)

data_example <- data.frame( x = c('Primary 1', 'Primary 2',
                                  'Secondary 1', 'Secondary 2'),
                            y = runif(4, 10, 100))

ggplot(data_example, aes(x,y, fill = x)) +
  geom_bar(stat = "identity") +
  labs(title = "Some Data (2011)",
       x = "Range", y = "Value") +
  dk_theme # <- Here's the theme!

```

### Statbank Demo
```{r get dataset code, echo = T}
library(dplyr)
dataset_names <- search_statbank_datasets("garda")
dataset_names[1,1]
dataset_names[1,2]
```


```{r use dataset code, echo = T}
industry_population <- get_cso_dataset(dataset_names$dataset_code[1])

industry_population %>% 
  filter(grepl("Both", Sex)) %>%
  filter(grepl("1991", `Census Year`)) %>% 
  filter(grepl("services", `Detailed Industrial Group`)) %>% 
  select(`Detailed Industrial Group`, `Census Year`, value) %>% 
  arrange(desc(value))


```

### Bringing It All Together
```{r altogether now, echo = T}

ggplot(industry_population %>% 
         filter(grepl("Garda|Defence", `Detailed Industrial Group`)) %>% 
         filter(grepl("Male|Female", Sex)) %>% 
         filter(!grepl("2016", `Census Year`)),
       aes(x = `Census Year`, y = value, fill = Sex)) + 
  facet_wrap(~`Detailed Industrial Group`, nrow = 1) +
  geom_bar(stat= "identity") +
  labs(title = "Population Working in Industrial Group", y = "Population") +
  dk_theme


```
